Evaluating the Performance and Energy Efficiency of COSMO-ART, a Coupled Numerical Weather Forecast and Chemical Transport Model

### 2014-04-09 trunk/paper_enahpc14/Sec2:Experimental-setup.tex
modifies "Evaluation" section; adds figure on T2S and E2S

\subsubsection{Monch (CSCS - ETH Zurich)}
Monch, which is installed  at the Swiss National Supercomputing Center
(CSCS)  of the ETH  Zurich, is  a 10  rack NEC-provided  (LX-2400) and
dual-socket Intel-based cluster, utilised from people that are part of
the   Swiss  Platform   for  Advanced   Scientific   Computing  (PASC,
\url{http://www.pasc-ch.org/}).   It  is   composed  of  312  standard
compute  nodes,  24  large-memory  compute nodes  and  24  huge-memory
compute nodes.   Each standard compute node comprises  2 Intel 10-core
Ivy Bridge EP E5-2660v2 processors operating at 2.2 GHz, offering 32GB
of DDR3 1600MHz  RAM, themselves connected by a  high speed Infiniband
network,  based on Mellanox  SX6036 managed  FDR (Fourteen  Data Rate)
switches, with a 56 Gb/s speed.

\subsubsection{Pilatus (CSCS - ETH Zurich)}
Pilatus is a dual-socket Intel  Sandy-Bridge based cluster used as Piz
Daint pre-post processing cluster.  It is composed of 42 compute nodes
with  a total  number of  1344 cores  in hyper-threading  and  2688 GB
aggregate memory and has 2  high-speed interconnects based on FDR: the
first is  dedicated to the MPI  traffic and the second  to the storage
high  speed traffic.  The  2 login  nodes  and the  42 computes  nodes
consists in  11 twin-pair Intel  E5-Series DALCO r2264i4t  2U scalable
compute  modules.  Each  module contains  4 compute  nodes based  on 2
Intel  8-core Intel  Xeon  E5-2670 processors  operating  at 2.6  GHz,
offering  64GB of  DDR3 1600MHz  RAM, themselves  connected by  a high
speed  Infiniband  network,  based  on  Mellanox  SX6036  managed  FDR
(Fourteen Data Rate) switches, with a 56 Gb/s speed.

@@

\subsubsection{CSCS - ETH Zurich}
E3METER  Intelligent Power Strips  (IPS) and  Monitors (IPM)  are high
quality electricity meters which enable to monitor and optimize energy
consumption of datacenters or large facilities.  A central data logger
- the  E3METER Data Concentrator  - aggregates  all measured  data and
makes them available in various forms (SNMP, Web, etc). Data transfers
between E3METER  IPS and E3METER Concentrator  use reliable narrowband
powerline  communication (PLC)  technology which  avoids the  need for
extra cabling.  It has  one internal temperature sensor. Two dedicated
extension  ports  can be  used  to  measure  external temperature  and
humidity  through E3METER  remote sensors.  It  features non-intrusive
voltage, current and frequency measurement through current transformer
with an accuracy of 1\%.

>>>>

\subsubsection{Monch (CSCS - ETH Zurich)}
Monch, which is installed  at the Swiss National Supercomputing Center
(CSCS)  of the ETH  Zurich, is  a 10  rack NEC-provided  (LX-2400) and
dual-socket Intel-based cluster, utilised from people that are part of
the   Swiss  Platform   for  Advanced   Scientific   Computing  (PASC,
\url{http://www.pasc-ch.org/}).   It  is   composed  of  312  standard
compute  nodes,  24  large-memory  compute nodes  and  24  huge-memory
compute nodes.   Each standard compute node comprises  2 Intel 10-core
Ivy Bridge  EP E5-2660v2 processors  operating at 2.2  GHz, themselves
connected by a high speed Infiniband network, based on Mellanox SX6036
managed  FDR switches,  with a  56  Gb/s speed.  Each core  has 32  KB
instruction and 32 KB  data L1 caches and 256 KB of  L2 cache. All the
10 core  share a 25.6 MB  L3 cache and  the platform has 32GB  of DDR3
1600MHz RAM.

\subsubsection{Pilatus (CSCS - ETH Zurich)}
Pilatus is a dual-socket Intel  Sandy-Bridge based cluster used as Piz
Daint pre-post processing cluster.  It is composed of 42 compute nodes
with  a total  number of  1344 cores  in hyper-threading  and  2688 GB
aggregate memory and has 2  high-speed interconnects based on FDR: the
first is  dedicated to the MPI  traffic and the second  to the storage
high  speed traffic.   The 2  login nodes  and the  42  computes nodes
consists in  11 twin-pair Intel  E5-Series DALCO r2264i4t  2U scalable
compute  modules.  Each  module contains  4 compute  nodes based  on 2
Intel  8-core Intel  Xeon  E5-2670 processors  operating  at 2.6  GHz,
themselves  connected by  a high  speed Infiniband  network,  based on
Mellanox SX6036 managed FDR switches,  with a 56 Gb/s speed. Each core
has  32 KB  instruction and  32 KB  data L1  caches and  256 KB  of L2
cache. All the 8 core share a 20 MB L3 cache and the platform has 64GB
of DDR3 1600MHz RAM.

@@

\subsubsection{CSCS - ETH Zurich}
Hardware platforms in our experiment include E3METER Intelligent Power
Strips  (IPS) and Monitors  (IPM) which  are high  quality electricity
meters  that enable  to  monitor and  optimize  energy consumption  of
datacenters or large facilities.  A  central data logger - the E3METER
Data  Concentrator  - aggregates  all  measured  data  and makes  them
available in  various forms (SNMP,  Web, etc). Data  transfers between
E3METER IPS and E3METER Concentrator use reliable narrowband powerline
communication  (PLC)  technology  which  avoids  the  need  for  extra
cabling.   It  has one  internal  temperature  sensor.  Two  dedicated
extension  ports  can be  used  to  measure  external temperature  and
humidity  through E3METER remote  sensors.  It  features non-intrusive
voltage, current and frequency measurement through current transformer
with an accuracy of 1\%.

### 2014-04-09 trunk/paper_enahpc14/paper.tex
adjusts structure of the document
modifies "Evaluation" section; adds figure on T2S and E2S

\author{Joseph~Charles  \and William~Sawyer  \and  Manuel~F.~Dolz \and
  Cristiano~Malossi}

@@

\institute{J.~Charles, W.~Sawyer 
	\at Swiss National Supercomputing Centre (CSCS)
	\\ CH-6900 Lugano, Switzerland  
	%\\ Tel: +41  (0) 91  610 8216
 	\\ E-mails:~{\{joseph.charles,william.sawyer\}@cscs.ch} 
        \and M.~F.~Dolz 
        \at Dept. of Informatics, University of Hamburg
	\\ D-22.527 Hamburg, Germany
        %\\ Tel: +49 (0) 40 460094-404
        \\ \email{manuel.dolz@informatik.uni-hamburg.de}  
        \and C.~Malossi  
        \at IBM Zurich Research Laboratory 
        \\  CH-8000 Zurich, Switzerland
        %\\  Tel: +41 (0) 44 724 8616
        \\ \email{ACM@zurich.ibm.com}}

@@

\begin{abstract}
  In this paper we present  COSMO-ART, an extension of the operational
  weather  forecast  model  of   the  German  weather  service  (DWD),
  developed for  the evaluation of the interactions  of reactive gases
  and aerosol particles  with the state of atmosphere  at the regional
  scale. It  includes secondary aerosols,  directly emitted components
  like  soot,  mineral  dust,  sea  salt and  biological  material  as
  pollen. Processes such  as emissions, coagulation, condensation, dry
  deposition,  wet removal,  and sedimentation  of aerosols  are taken
  into account.   The overall performance  of this application  on HPC
  systems is analysed  by a profiling study to  determine hotspots and
  identify critical paths.   Moreover, we describe measurement devices
  and  energy-aware   techniques  employed  to   evaluate  the  energy
  footprint of the considered application and to get detailed insights
  about power bottlenecks.  Our motivation is to improve corresponding
  code   sections  to  sustain   high  performance   while  minimizing
  energy-to-solution.   This  preliminary  work  sets  the  basis  for
  subsequent studies to tackle  challenges related to energy efficient
  high performance computing in the framework of the Exa2Green project
  (\url{http://exa2green.eu/}).

\keywords{High performance computing  \and Energy-aware computing \and
  Green computing  \and Numerical weather  prediction \and Atmospheric
  chemistry  \and  Aerosols  modelling  \and  Profiling  methods  \and
  Benchmark analysis \and COSMO-ART coupled model}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

\section{Power-performance measurement framework}
\label{sec:2}
\input{Sec2:Power-measurement.tex}

@@

\begin{acknowledgements}
This  research  was  supported   by  the  Exa2Green  research  project
co-funded  under the  EU  7th Framework  Program  Future and  Emerging
Technologies (FET) Proactive Initiative: Minimising Energy Consumption
of Computing to  the Limit (MINECC).  The authors  would like to thank
the  High  Performance  and  High  Productivity  Computing  Initiative
(\url{www.hp2c.ch}) for  results that will be  leveraged in subsequent
code refactoring.
\end{acknowledgements}

>>>>

\author{Joseph~Charles  \and William~Sawyer  \and  Manuel~F.~Dolz \and
  Sandra~Catalan}

@@

\institute{J.~Charles, W.~Sawyer 
	\at Swiss National Supercomputing Centre (CSCS)
	\\ CH-6900 Lugano, Switzerland  
 	\\ E-mails:~{\{joseph.charles,william.sawyer\}@cscs.ch} 
        \and M.~F.~Dolz 
        \at Dept. of Informatics, University of Hamburg (UHAM)
	\\ DE-22527 Hamburg, Germany
        \\ \email{manuel.dolz@informatik.uni-hamburg.de}  
        \and S.~Catalan
        \at Jaume I University of Castell\'on (UJI)
	\\ ES-12071 Castell\'on, Spain
        \\ \email{catalans@icc.uji.es}}

@@

\begin{abstract}
  In this paper we present  COSMO-ART, an extension of the operational
  weather  forecast  model  of   the  German  weather  service  (DWD),
  developed for  the evaluation of the interactions  of reactive gases
  and aerosol particles  with the state of atmosphere  at the regional
  scale. It  includes secondary aerosols,  directly emitted components
  like  soot,  mineral  dust,  sea  salt and  biological  material  as
  pollen. Processes such  as emissions, coagulation, condensation, dry
  deposition,  wet removal,  and sedimentation  of aerosols  are taken
  into account.   The overall performance  of this application  on HPC
  systems is analysed  by a profiling study to  determine hotspots and
  identify  critical  CPU paths.   Moreover,  we describe  measurement
  devices and energy-aware techniques  employed to evaluate the energy
  footprint of the considered application and to get detailed insights
  about power bottlenecks.  Our motivation is to improve corresponding
  code   sections  to  sustain   high  performance   while  minimizing
  energy-to-solution.   This  preliminary  work  sets  the  basis  for
  subsequent studies to tackle  challenges related to energy efficient
  high performance computing in the framework of the Exa2Green project
  (\url{http://exa2green.eu/}).

\keywords{High performance computing  \and Energy-aware computing \and
  Green computing  \and Numerical weather  prediction \and Atmospheric
  chemistry  \and  Aerosols  modelling  \and  Profiling  methods  \and
  Benchmark    analysis   \and    COSMO-ART    coupled   model    \and
  Time-to-solution \and Energy-to-solution \and Manycore}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

\section{Experimental setup}
\label{sec:2}
\input{Sec2:Experimental-setup.tex}

@@

\begin{acknowledgements}
This  research  was  supported   by  the  Exa2Green  research  project
co-funded  under the  EU  7th Framework  Program  Future and  Emerging
Technologies (FET) Proactive Initiative: Minimising Energy Consumption
of Computing  to the Limit  (MINECC).  We also  gratefully acknowledge
the  High  Performance  and  High  Productivity  Computing  Initiative
(\url{www.hp2c.ch}) for  results that will be  leveraged in subsequent
code refactoring.
\end{acknowledgements}
